Compositional Equivalences Based on Open pNets

### 2022-10-18 2016-Weak-FH-Bisim/WeakBisim.tex
corrections

Bisimulation over a symbolic and open model like open pNets or open automata is  different from the classical notion of bisimulation because it cannot rely on the equality over a finite set of action labels. Classical bisimulations require to exhibit, for each transition of one system, a transition of the other system that simulates it. Instead, bisimulation for open automata  relies on the simulation of each open transition of one automaton by \emph{a set of} open transitions of the other one, that should cover all the 
cases where the original transition can be triggered. %\TODO{Eric: is this like in symbolic bisimulation ?}
\ERIC{This is similar to the early and late symbolic bisimulation equivalences for value-passing CCS \cite{HENNESSY:1995}, though we need more general definitions in our setting.}

Compositionality of bisimilarity in our model comes from the specification of the interactions, including actions of the holes. 
\ERIC{This is quite different from the works on contextual equivalences, e.g. \cite{Larsen:91}; we will give more details on this in section \ref{sec:RW}.}
In pNets, synchronisation vectors define the possible interactions between the pNet that fills the hole and the surrounding pNets. In open automata, this is reflected by symbolic hypotheses that depend on the actions of the holes. This additional specification is the price to pay to obtain the compositionality of bisimilarity that cannot be guaranteed in traditional process algebras.

@@

The semantics of open pNets will be defined  as an open automaton. An open
automaton is an automaton where each transition composes transitions of several LTSs with
action of some holes, the transition occurs if some predicates hold, and can involve a 
set of state modifications. This section defines open automata and a bisimulation theory for them. This section is an improved version of the formalism described in \cite{henrio:Forte2016}, extending the automata with a notion of global variable, which makes the state of the automaton more explicit. We also adopt a semantics and logical interpretation of the automata that intuitively can be stated as follows: ``if a transition belongs to an open automaton, any refinement of this transition also belongs to the automaton''. \LUDO{Our open automata are clearly inspired by the work  of de Simone on formatting of SOS rules~\cite{deSimone85}. A precise comparison with related works can be found in Section~\ref{sec:RW}.}

@@

\begin{thm}[Congruence]\label{thm-congr-eq}
	Consider an open pNet:
	$\pNet = \mylangle \pNet_i^{i\in I}, \Sort_j^{j\in J}, 
	\set{\symb{SV}}\myrangle$.
	Let $j_0\in J$ be a hole. Let $\pNetQ$ and $\pNetQ'$ be two FH-bisimilar pNets such that\footnote{Note that $\Sortop(\pNetQ)=\Sortop(\pNetQ')$ is 
	ensured by 
	strong bisimilarity.} 
	$\Sortop(\pNetQ)=\Sortop(\pNetQ')=\Sort_{j_0}$. Then 
	$\pNet[\pNetQ]_{j_0}$ and 
	$\pNet[\pNetQ']_{j_0}$ are FH-bisimilar.
\end{thm}

@@

There is also a clear relation with the seminal works on rule formats for Structured Operational Semantics, e.g. DeSimone format, GSOS, and conditional rules with or without negative premises \cite{deSimone85,Bloom:1995,GROOTE1992202,VANGLABBEEK2004229}.  The Open pNets model provides a way to define operators similar to these rules formats, but with quite different aim and approach. A formal comparison would be interesting, though not trivial. What we can say easily is that: the pNet format syntactically encompasses  De Simone, GSOS, and conditional premises rules. Then our compositionality result is more powerful than their classical results, but this is not a surprise, as we rely on a (sufficient) syntactic hypothesis on a particular system, rather than the general rules defining an operator.  Last, we intentionally do not accept negative premises, that would be more  to put into practice in our implementation. This extension could be studied in future work.

>>>>

Bisimulation over a symbolic and open model like open pNets or open automata is  different from the classical notion of bisimulation because it cannot rely on the equality over a finite set of action labels. Classical bisimulations require to exhibit, for each transition of one system, a transition of the other system that simulates it. Instead, bisimulation for open automata  relies on the simulation of each open transition of one automaton by \emph{a set of} open transitions of the other one, that should cover all the 
cases where the original transition can be triggered. %\TODO{Eric: is this like in symbolic bisimulation ?}
\ERIC{This is similar to the early and late symbolic bisimulation equivalences for value-passing CCS \cite{HENNESSY:1995}, though we use more general definitions in our setting.}

Compositionality of bisimilarity in our model comes from the specification of the interactions, including actions of the holes. 
\ERIC{This is quite different from the works on contextual equivalences, e.g. \cite{Larsen87, Larsen:91}; we will give more details on this in section \ref{sec:RW}.}
In pNets, synchronisation vectors define the possible interactions between the pNet that fills the hole and the surrounding pNets. In open automata, this is reflected by symbolic hypotheses that depend on the actions of the holes. This additional specification is the price to pay to obtain the compositionality of bisimilarity that cannot be guaranteed in traditional process algebras.

@@

The semantics of open pNets will be defined  as an open automaton. An open
automaton is an automaton where each transition composes transitions of several LTSs with
action of some holes, the transition occurs if some predicates hold, and can involve a 
set of state modifications. This section defines open automata and a bisimulation theory for them. This section is an improved version of the formalism described in \cite{henrio:Forte2016}, extending the automata with a notion of global variable, which makes the state of the automaton more explicit. We also adopt a semantics and logical interpretation of the automata that intuitively can be stated as follows: ``if a transition belongs to an open automaton, any refinement of this transition also belongs to the automaton''. \LUDO{Our open automata are clearly inspired by the work  of De Simone on formatting of SOS rules~\cite{deSimone85}. A precise comparison with related works can be found in Section~\ref{sec:RW}.}

@@

\begin{thm}[Congruence]\label{thm-congr-eq}
	Consider an open pNet \\
	$\pNet = \mylangle \pNet_i^{i\in I}, \Sort_j^{j\in J}, 
	\set{\symb{SV}}\myrangle$.
	Let $j_0\in J$ be a hole. Let $\pNetQ$ and $\pNetQ'$ be two FH-bisimilar pNets such that\footnote{Note that $\Sortop(\pNetQ)=\Sortop(\pNetQ')$ is 
	ensured by 
	strong bisimilarity.} 
	$\Sortop(\pNetQ)=\Sortop(\pNetQ')=\Sort_{j_0}$. Then 
	$\pNet[\pNetQ]_{j_0}$ and 
	$\pNet[\pNetQ']_{j_0}$ are FH-bisimilar.
\end{thm}

@@

There is also a clear relation with the seminal works on rule formats for Structured Operational Semantics, e.g. De Simone format, GSOS, and conditional rules with or without negative premises \cite{deSimone85,Bloom:1995,GROOTE1992202,VANGLABBEEK2004229}.  The Open pNets model provides a way to define operators similar to these rules formats, but with quite different aim and approach. A formal comparison would be interesting, though not trivial. What we can say easily is that: the pNet format syntactically encompasses  De Simone, GSOS, and conditional premises rules. Then our compositionality result is more powerful than their classical results, but this is not a surprise, as we rely on a (sufficient) syntactic hypothesis on a particular system, rather than the general rules defining an operator.  Last, we intentionally do not accept negative premises, that would be more  to put into practice in our implementation. This extension could be studied in future work.
